Reasoning Graphically about Quantum Computation

### 2008-02-25 papers/reasoning-graphically/paper.tex
working...
git-svn-id: https://isaplanner.svn.sourceforge.net/svnroot/isaplanner/trunk/quantomatic@1370 f8f375fb-4f10-0410-8d47-b6fbdc7ca053

@@

@@

!-Boxes

In graph patterns, variable nodes generalise the concept of a
half-edge. Variable nodes, unlike boundary-nodes in the concrete
graph, are not classified into domain and co-domain. The intuition of
a variable node is that it can be replaced by any concrete node. Thus,
unlike nodes in conrete a graph, variable nodes do not define an
ordering on the edges that are incident. Instead, the edges that touch
a variable node are simply a set.

The type of a variable nodes is thus defined as a multiset rather than
a tensor. Any tensor type can be lifted into a multiset type in the
obvious manner, which we denote by the operation {\tt
  mset\_of\_tensor}. For example, {\tt mset\_of\_tensor($a \tensor b
  \tensor a$) = $\{(a,2), (b,1)\}$}. We say that a tensor-type, $T$,
is an instance of a multiset type, $M$, when it's lifting into a
multiset is a super-set of $M$: $ M \subseteq
\mathtt{mset\_of\_tensor}(T) \Longrightarrow T \in M$. The intuition
here is that an instance of a variable node is allowed to have as many
extra edges as it likes as long as it at least has the ones mentioned
in the original graph. A variable node, $a$ is be said to an instance
of another variable node $b$ when the multiset representing the type of
$a$ is a superset of the multiset for the type of $b$. Thus, with
regard to the underlying semantics, $a$ being a subtype of $b$ means
that the set of instances of $a$ are a subset of the instances of $b$.

\begin{figure}[t]
%  \scalebox{1.0}{\includegraphics{images/node-var-instance.eps}}
*** Add figure *** 
\label{node-variable-instances-fig}\caption{The concrete graphs $a$,
  $b$, and $c$ are instances of $x$, a graph with variables nodes.}
\end{figure}

A graph with variable nodes is given a formal semantics by being
interpretated as a set of concrete graphs. We refer members of
interpretation set as instances. Formally, an instance of a graph with
variable-nodes is the graph where every variable node has been
replaced by some instance of it. For example, in
Figure~\ref{node-variable-instances-fig}, the graphs $a$, $b$ and $c$
are instances of the graph $x$. Observe that when a variable-node has
degree one, then it can be instantiated as a boundary edge of either
the domain or co-domain.

\subsection{Compositonality}

In this section, we define a compositonality principle for graphs
patterns, that is for graphs with variable-nodes and !-boxes. This is
related to compositionality of the underlying concrete graphs by the
following instantiation homomorphism:

$\llbracket G \cric H \rrbracket = \llbracket G \rrbracket \cric \llbracket H \rrbracket$

>>>>

\newcommand{\vinterp}[1]{\lfloor\hspace{-0.27em}|\, #1\, |\hspace{-0.27em}\rfloor}
\newcommand{\binterp}[1]{\lceil\hspace{-0.27em}|\, #1\, |\hspace{-0.27em}\rceil}
\newcommand{\minterp}[1]{\llbracket #1 \rrbracket}
\newcommand{\sinterp}[1]{| #1 |}

@@

In graph patterns, variable nodes generalise the concept of a
boundary node. Variable nodes, unlike boundary-nodes in the concrete
graph, are not classified into domain and co-domain. The intuition of
a variable node is that it can be replaced by any concrete node. Thus,
unlike nodes in conrete a graph, variable nodes do not define an
ordering on the edges that are incident. Instead, the edges that touch
a variable node are simply a set.

The type of a variable nodes is thus defined as a multiset rather than
a tensor. Any tensor type can be lifted to a multiset type in the
obvious manner, which we denote by the operation {\tt
  mset\_of\_tensor}. For example, {\tt mset\_of\_tensor($a \tensor b
  \tensor a$) = $\{(a,2), (b,1)\}$}. We say that a tensor-type, $T$,
is an instance of a multiset type, $M$, when it's lifting into a
multiset is a super-set of $M$: $ M \subseteq
\mathtt{mset\_of\_tensor}(T) \Longrightarrow T \in M$. The intuition
here is that an instance of a variable node is allowed to have as many
extra edges as it likes as long as it at least has the ones mentioned
in the original graph. A variable node, $a$ is be said to an instance
of another variable node $b$ when the multiset representing the type of
$a$ is a superset of the multiset for the type of $b$. Thus, with
regard to the underlying semantics, $a$ being a subtype of $b$ means
that the set of instances of $a$ are a subset of the instances of $b$.

\begin{figure}[t]
%  \scalebox{1.0}{\includegraphics{images/node-var-instance.eps}}
*** Add figure *** 
\label{node-variable-instances-fig}\caption{The concrete graphs $a$,
  $b$, and $c$ are instances of $x$, a graph with variable nodes.}
\end{figure}

A graph with variable nodes, $G$, can be given a formal semantics by
being interpretated as a set of concrete graphs which is denoted by
$\vinterp{G}$. We refer to members of this set as
instances. Formally, an instance of a graph with variable nodes is the
graph where every variable node has been replaced by some instance of
it. For example, in Figure~\ref{node-variable-instances-fig}, the
graphs $a$, $b$ and $c$ are instances of the graph $x$. Observe that
when a variable node has degree one, then it can be instantiated as a
boundary node in either the domain or co-domain.

\subsection{Plugging}

@@

The set of concrete graphs for a graph $G$ with !-boxes, is denoted by
$\binterp{G}$.

\subsection{Matching}

The purpose of our graph patterns is to express and formalise, in a
declarative manner, the intuitive idea of graph matching for graphs
that represent Compact Closed Categories. In the previous sections we
defined ways to extend the concrete graph representation with variable
nodes and !-boxes. In this section we combine these to first define
the matching relation between graph patterns and concrete graphs and
then introduce a matching relation between graph patterns.

\subsubsection{Concrete Matches}

We give a semantics to these in terms of a set of concrete graphs,
denoted by $\minterp{G}$, and say that a member of this set is
\emph{matched} by the graph pattern $G$. Formally, the set of matches
are defined as follows:

$\minterp{G} = \sinterp{\vinterp{\binterp{G}}}$

where $\sinterp{G}$ represents the set of graphs that strictly contain
the graph $G$.

\subsubsection{Pattern Matches}

In this section, we define a compositonality principle for graphs
patterns, that is for graphs with variable nodes and !-boxes. This is
related to compositionality of the underlying concrete graphs by the
following instantiation homomorphism:

$\llbracket G \circ H \rrbracket = \llbracket G \rrbracket \circ \llbracket H \rrbracket$
